### 2015-05-12 intro.tex
added intro to the implementation

\whoop currently supports only Linux drivers, but the approach is generic enough and thus we believe that it can be used to analyze drivers from other operating systems (e.g.\ Windows) as well as other concurrent programs with similar structure (e.g.\ filesystems).

>>>>

\whoop currently supports only Linux drivers, but the approach is generic, and thus we believe that it can be used to analyze drivers from other operating systems (e.g.\ Windows) as well as other concurrent programs that use a similar programming model (e.g.\ filesystems).

### 2015-05-12 whoop.tex
added intro to the implementation

We now present \emph{symbolic pairwise lockset analysis}, a novel method for data race analysis in device drivers.  In Section~\ref{sec:symbolicpairwise} we describe how the approach works in a semi-formal manner, with respect to a simple concurrent programming model.  In Section~\ref{sec:whoopimplementation} we explain how we have implemented symbolic pairwise lockset analysis in a practical tool, \whoop, that can be applied directly to driver source code, to check whether pairs of driver entry points can race with one another.

@@

\subsection{Symbolic Pairwise Lockset Analysis}\label{sec:symbolicpairwise}

@@

\medskip\noindent\textbf{Concurrent Programming Model }
%
We consider a simple concurrent programming model in which an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  In the context of device drivers, a thread template corresponds to a driver entry point, and multiple instances of the same thread template may execute concurrently, just as multiple invocations of a single driver entry point may be concurrent.  Further threads may be launched at any point during execution; in the context of device drivers this corresponds to the OS invoking additional driver entry points.  For ease of presentation only, our model does not feature aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in Section~\ref{sec:whoopimplementation} we discuss interesting practical issues arising from the handling of a full-blown language.

@@

\subsection{Implementation in \whoop}\label{sec:whoopimplementation}

\ADComment{Acknowledge very clearly that the simple concurrent programming model of Section~\ref{...} is deliberately idealistic.  In practice, we do not know upfront what the entry points for a driver are, we don't have named set of locks, and rather than a given set of named shared variables, we have arbitrary memory accesses via pointers.  We now explain how we have taken the conceptual ideas from Section~\ref{...} and used them to build a practical tool.}

>>>>

We now present \emph{symbolic pairwise lockset analysis}, a novel method for data race analysis in device drivers.  In Section~\ref{sec:symbolicpairwise} we describe how the approach works in a semi-formal manner, with respect to a simple concurrent programming model.  In Section~\ref{sec:implementation} we explain how we have implemented symbolic pairwise lockset analysis in a practical tool, \whoop, that can be applied directly to driver source code, to check whether pairs of driver entry points can race with one another.

@@

\subsection{Symbolic Pairwise Lockset Analysis}
\label{sec:symbolicpairwise}

@@

\medskip\noindent\textbf{Concurrent Programming Model }
%
We consider a simple concurrent programming model in which an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  In the context of device drivers, a thread template corresponds to a driver entry point, and multiple instances of the same thread template may execute concurrently, just as multiple invocations of a single driver entry point may be concurrent.  Further threads may be launched at any point during execution; in the context of device drivers this corresponds to the OS invoking additional driver entry points.  For ease of presentation only, our model does not feature aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in Section~\ref{sec:implementation} we discuss interesting practical issues arising from the handling of a full-blown language.

@@

\subsection{Implementation in \whoop}
\label{sec:implementation}

The simple concurrent programming model discussed in Section~\ref{sec:symbolicpairwise} is deliberately idealistic to describe our symbolic verification technique. In practice, the Linux drivers we analyze are written in the C language, we do not know upfront which are the entry points for a driver, we do not have a named set of locks, and rather than having a given set of named shared variables, we have arbitrary memory accesses via pointers. We now explain how we have taken the conceptual ideas from Section~\ref{sec:symbolicpairwise} and used them to build a practical tool for soundly detecting data races in drivers.
